XOR-Some Prolific Pilot 2

### 2021-04-09 analysis/xor-some-prolific-pilot2.Rmd
knit

Plot inference strength ratings against raw predictor ratings for each participant / each item:
```{r}
d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

@@

Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bard indicate the standard deviation:
```{r, fig.height=4, fig.width=8}
d_critical_predictorsXtarget_summary <- d_critical_predictorsXtarget %>% 
  group_by(title, predictors, main_type) %>%
  summarise(mean_target = mean(target),
            mean_predictor = mean(predictor_response), 
            sd_target = sd(target),
            sd_predictor = sd(predictor_response))

@@

Plot z-scored predictor ratings against z-scored target ratings
```{r}
d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

@@

## Stats
Maximal models on raw ratings:
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs
model_xor <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

@@

Exploring plots of posterior samples for pairwise parameters shows that there is a visible correlation between prior & competence, and a somewhat weaker correlation between competence & relevance. The interaction estimates also show a correlation with the main effects (as a rough overall judgment).   
```{r}
shinystan::launch_shinystan(model_xor)
```
Plot correlations of main effects
```{r}
bayesplot::mcmc_pairs(model_xor, pars = c("b_prior", "b_competence", "b_relevance"))
```

@@

For "some", the correlation of prior and relevance is higher than for prior and competence; there is also a visible correlation of competence and relevance. 
```{r}
shinystan::launch_shinystan(model_some)
```
```{r}
bayesplot::mcmc_pairs(model_some, pars = c("b_prior", "b_competence", "b_relevance"))
```

#### Smaller models with raw data to check for colinearity 
Models on raw data with only one predictor: 
```{r, results='hide'}
# xor
model_xor_prior <- brm(
  target ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

@@

#### Stats on z-scored data
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs on z-scored data
model_xor_zScored <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

@@

In summary, the different models provide somewhat different results, specially when using raw ratings vs. binary predictors.

>>>>

Plot inference strength ratings against raw predictor ratings from each participant across items:
```{r}
d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

@@

Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bars indicate the standard deviation of the mean:
```{r, fig.height=4, fig.width=8}
d_critical_predictorsXtarget_summary <- d_critical_predictorsXtarget %>% 
  group_by(title, predictors, main_type) %>%
  summarise(mean_target = mean(target),
            mean_predictor = mean(predictor_response), 
            sd_target = sd(target),
            sd_predictor = sd(predictor_response))

@@

Plot z-scored predictor ratings against z-scored target inference ratings
```{r}
d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

@@

## Stats

Maximal models on raw ratings:
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs
model_xor <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

@@

Exploring plots of posterior samples for pairwise parameters shows that there is a visible correlation between prior & competence, and a somewhat weaker correlation between competence & relevance. The interaction estimates also show a correlation with the main effects (as a rough overall judgment).   
```{r}
#shinystan::launch_shinystan(model_xor)
```
Plot correlations of main effects
```{r}
bayesplot::mcmc_pairs(model_xor, pars = c("b_prior", "b_competence", "b_relevance"))
```

@@

For "some", the correlation of prior and relevance is higher than for prior and competence; there is also a visible correlation of competence and relevance. 
```{r}
#shinystan::launch_shinystan(model_some)
```
```{r}
bayesplot::mcmc_pairs(model_some, pars = c("b_prior", "b_competence", "b_relevance"))
```

#### Smaller models with raw data to check for collinearity 
Models on raw data with only one predictor: 
```{r, results='hide'}
# xor
model_xor_prior <- brm(
  target ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

@@

#### Stats on z-scored data
The same stats (full, one predictor, to predictors) on z-scored data:
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs on z-scored data
model_xor_zScored <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

@@
